---
title: "HR Attrition Predictions"
author: "Nicole Bartholow/Chase Henderson"
date: "12/2/2018"
output: pdf_document
---

```{r Setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(DataExplorer)
library(plyr)
library(dplyr)
library(ggplot2)
library(kableExtra)
library(openintro)
library(reshape2)
library(gridExtra)
library(grid)
library(mlr)
library(caret)
library(e1071)
library(AppliedPredictiveModeling)
library(GGally)
library(corrplot)
```

#Introduction
###Talk about merger with Sierra

</p>


####First we bring in the data provided and clean it up for analysis.  
```{r Read Attrition Data Files}
#set working directory for local computer
#REPRODUCIBLE TIP - CHANGE THIS WORKING DIRECTORY FOR YOUR REPRODUCTION ENVIRONMENT
#set working directory for local computer
setwd("/Users/nicolealphin/Desktop/DataScience/MSDS6306-DoingDataScience/HW/CaseStudy2")
DFTrain <- read.csv("./Analysis/Data/CaseStudy2-data.csv", header=TRUE) #import HR Data into DFTrain dataframe
DFTest <- read.csv("./Analysis/Data/CaseStudy2Validation.csv", header=TRUE) #import HR Validation Set into DFVal dataframe
```

#Remove columns that are irrelevant or have only one value
#Explain the columns that are being removed and why
```{r remove unneeded columns}
DFTrain<- subset(DFTrain, select=-c(ID, EmployeeCount,EmployeeNumber,Over18,StandardHours, Rand ))
DFTest<- subset(DFTest, select=-c(ID, EmployeeCount,EmployeeNumber,Over18,StandardHours, Rand ))
```

#Analyze current population
#Explain Happiness Index
#Correlation Plots
#Not surprisingly, all the Years and Age are correlated
#Surprisingly, the Satisfactions do not show any correlation - that's why they should be added
```{r analyse satisfactions and attrition by role and level}
DFTrain$Happiness <- DFTrain$EnvironmentSatisfaction+DFTrain$JobSatisfaction+DFTrain$RelationshipSatisfaction+DFTrain$WorkLifeBalance 
DFTest$Happiness <- DFTest$EnvironmentSatisfaction+DFTest$JobSatisfaction+DFTest$RelationshipSatisfaction+DFTest$WorkLifeBalance 

#Put Datasets Together for complete counts and pictures.  I'm including these in the entire picture with Attrition information because we have it.
DFTotal <- rbind(DFTrain, DFTest)

DFTrain$EducationField <- as.integer(DFTrain$EducationField)
DFTrain$JobRole <- as.integer(DFTrain$JobRole)
DFTrain$OverTime <- as.integer(DFTrain$OverTime)
DFTrain$MaritalStatus <- as.integer(DFTrain$MaritalStatus)
DFTrain$BusinessTravel <- as.integer(DFTrain$BusinessTravel)
DFTrain %>%
  filter(Attrition == "Yes") %>%
  select_if(is.numeric) %>%
  cor() %>%
  corrplot::corrplot()
DFTrain %>%
  filter(Attrition == "No") %>%
  select_if(is.numeric) %>%
  cor() %>%
  corrplot::corrplot()

#Create barplot for Department counts
#x must be continuous here, try again
p <- ggplot(DFTotal, aes(x = Department))+geom_bar(fill="#3182bd", colour="black")
p + scale_fill_brewer(palette = "Blues") + 
  ggtitle("Distribution of Sierra Employees by Department")+ 
  theme( plot.title = element_text(hjust = 0.5))

#Create a table of JobRoles by level
#RoleLevelCount <- data.frame(table(DFTrain$JobRole, DFTrain$JobLevel), decreasing=TRUE)
RoleLevelCount <- data.frame(table(DFTotal$JobRole, DFTotal$JobLevel))
colnames(RoleLevelCount)<-c("JobRole", "JobLevel", "Freq")#set column names to improve readability
RoleLevelCount
RoleLevelCount <- dcast(RoleLevelCount, JobRole ~ JobLevel)
#RoleLevelCount%>%
#  arrange_at(2:5, desc)

colnames(RoleLevelCount) <- c("JobRole", "Level 1", "Level 2", "Level 3", "Level 4", "Level 5")
RoleLevelCount$Total <- RoleLevelCount$`Level 1`+RoleLevelCount$`Level 2`+RoleLevelCount$`Level 3`+RoleLevelCount$`Level 4`+RoleLevelCount$`Level 5`

RoleLevelCount%>%
  arrange(match(JobRole, c("Sales Representative", "Human Resources", "Laboratory Technician", "Research Scientist", "Sales Executive", "Manufacturing Director", "Healthcare Representative", "Research Director", "Manager")))%>%
  kable() %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")


#Lets just do a straight stacked graph without proportions on JobRole
RoleAttritionCount <- data.frame(table(DFTotal$JobRole, DFTotal$Attrition))
colnames(RoleAttritionCount)<- c("JobRole", "Attrition","NumberOfEmployees")
p <- ggplot(RoleAttritionCount, aes(x=reorder(JobRole, -NumberOfEmployees), y=NumberOfEmployees, fill=Attrition))+geom_bar(stat="identity", colour="black")
p + scale_fill_manual(values=c("#9ecae1", "#3182bd"))+
  ggtitle("Employee Population by Job Role with Attrition Count")+ 
  theme( axis.text.x=element_text(angle = 90,vjust = 0), plot.title = element_text(hjust = 0.5))+ 
  xlab("JobRoles") + ylab("Number of Employees") 


#Let's get proportional stacked graphs to work!!!
#first you need a summarized view of the data
#RoleCount <- data.frame(sort(table(DFTotal$JobRole))) 
RoleLevelCount <- data.frame(table(DFTotal$JobRole, DFTotal$JobLevel))
#create a summary table for Attrtion by JobRole
RoleAttritionCount <- data.frame(table(DFTotal$JobRole, DFTotal$Attrition))
#Calculate percent attrition by role and add it as a new column
RA <- ddply(RoleAttritionCount, "Var1", transform,
            percent_attrition = Freq / sum(Freq) * 100)
colnames(RA)<-c("JobRole", "Attrition", "Freq", "PercentAttrition")#set column names to improve readability
RA$Order<- c(13, 14, 5, 6, 3, 4, 15, 16, 11, 12, 17, 18, 9, 10, 7, 8, 1, 2)
RA <- RA[order(RA$Order),]
rownames(RA) <- 1:nrow(RA)

#Proportional Stacked Bar Graph of Attrition by Job Role
ggplot(RA, aes(x=reorder(JobRole, Order), y=PercentAttrition, fill=Attrition)) +
    geom_bar(stat="identity", colour="black") +
    guides(fill=guide_legend(reverse=FALSE)) + 
    theme(axis.text.x=element_text(angle = 90,vjust = 0), plot.title = element_text(hjust = 0.5))+ 
    ggtitle("Percentage Attrition by JobRole")+
  scale_fill_manual(values=c("#9ecae1", "#3182bd"))+ 
    xlab("Job Roles") + ylab("Percent Attrition") 

#scale_fill_manual(values=c("#9ecae1", "#3182bd"))+ 
#scale_fill_manual(values=wes_palette(n=3, name="Zissou1"))

#Now let's look at that by Levels
DFTotal$JobLevel <- as.factor(DFTotal$JobLevel)
LevelAttritionCount <- data.frame(table(DFTotal$JobLevel, DFTotal$Attrition))
colnames(LevelAttritionCount)<- c("JobLevel", "Attrition","NumberOfEmployees")
p <- ggplot(LevelAttritionCount, aes(x=reorder(JobLevel, -NumberOfEmployees), y=NumberOfEmployees, fill=Attrition))+geom_bar(stat="identity", colour="black")
p + scale_fill_manual(values=c("#9ecae1", "#3182bd"))+
  ggtitle("Employee Population by Job Level with Attrition Count")+ 
  theme( axis.text.x=element_text(angle = 90,vjust = 0), plot.title = element_text(hjust = 0.5))+ 
  xlab("Job Levels") + ylab("Number of Employees") 

LA <- ddply(LevelAttritionCount, "JobLevel", transform,
            percent_attrition = NumberOfEmployees / sum(NumberOfEmployees) * 100)
colnames(LA)<-c("JobLevel", "Attrition", "Freq", "PercentAttrition")#set column names to improve readability

#Proportional Stacked Bar Graph of Attrition by Job Level
ggplot(LA, aes(x=JobLevel, y=PercentAttrition, fill=Attrition)) +
  geom_bar(stat="identity", colour="black") +
  guides(fill=guide_legend(reverse=FALSE)) + 
  theme(axis.text.x=element_text(angle = 90,vjust = 0), plot.title = element_text(hjust = 0.5))+ 
  ggtitle("Percentage Attrition by JobLevel")+
  scale_fill_manual(values=c("#9ecae1", "#3182bd"))+ 
  xlab("Job Levels") + ylab("Percent Attrition") 


#Plot of Happiness for Entire Company by Job Role
qplot(JobRole, Happiness, geom="boxplot", data=DFTotal)+ 
  theme(axis.text.x=element_text(angle = 90,vjust = 0), plot.title = element_text(hjust = 0.5))+ 
  ggtitle("Cumulative Happiness Score Across Departments")+
  scale_fill_manual(values=c("#9ecae1", "#3182bd"))

bp<-ggplot2::ggplot(DFTotal, aes(x=JobRole, y=Happiness)) +
  geom_boxplot()+ 
  theme(axis.text.x=element_text(angle = 90,vjust = 0), plot.title = element_text(hjust = 0.5))+ 
  ggtitle("Cumulative Happiness Score Across Departments")+
  scale_fill_manual(values="3182bd")
bp
```

```{r Correlations between Attrition and other variables }
DFTrain <- transform(DFTrain, Year1= ifelse(YearsAtCompany== 1, 1, 2))
DFTest <- transform(DFTest, Year1= ifelse(YearsAtCompany== 1, 1, 2))
DFTrain$MSInt <- as.integer(DFTrain$MaritalStatus)
DFTrain <- transform(DFTrain, Single= ifelse(MSInt== 1, 1, 2))
DFTest$MSInt <- as.integer(DFTest$MaritalStatus)
DFTest <- transform(DFTest, Single= ifelse(MSInt== 1, 1, 2))
#Correlations between Attrition and Predictors 
#Attrtion and Overtime- 0.246
DFTrain$Attrition<-as.integer(DFTrain$Attrition)
DFTotal$OverTime<-as.integer(DFTotal$OverTime)
AttOT <-cor(DFTotal$Attrition, DFTotal$OverTime, use = "complete.obs") 

#Attrition and Time with Company <2 years - AttYears = -0.181
AttYears <-cor(DFTrain$Attrition, DFTrain$Year1, use = "complete.obs") 

#Attrition and  Marital Satus - 0.162
DFTrain$MaritalStatus<-as.integer(DFTrain$MaritalStatus)
AttMS <-cor(DFTrain$Attrition, DFTrain$MaritalStatus, use = "complete.obs") 
#Derived Marital Status
AttMSI <-cor(DFTrain$Attrition, DFTrain$MSInt, use = "complete.obs") # - 0.18

#Attrition and NCW - AttJobInvolvement = -0.150
AttJI <-cor(DFTrain$Attrition, DFTrain$JobInvolvement, use = "complete.obs") 

#Attrition and EnvSat - AttES = -0.126
AttES <-cor(DFTrain$Attrition, DFTrain$EnvironmentSatisfaction, use = "complete.obs") 
```

##Prediction Models
#kNN Prediction Model
```{r kNN Prediction Model}
DFTrak <- DFTrain
DFTrak$Attrition <- as.integer(DFTrak$Attrition)
DFTrak$MaritalStatus <- as.integer(DFTrak$MaritalStatus)
DFTrak$OverTime <- as.integer(DFTrak$OverTime)

DFTek <- DFTest
DFTek$Attrition <- as.integer(DFTek$Attrition)
DFTek$MaritalStatus <- as.integer(DFTek$MaritalStatus)
DFTek$OverTime <- as.integer(DFTek$OverTime)

results1 <- class::knn(DFTrak[,c(16, 20, 13)], DFTek[,c(16,20, 13)], DFTrak$Attrition, k=7)
DFTek$AttPred1 <- results1
kNN_Prediction <- results1
confusionMatrix(table(DFTek$Attrition,DFTek$AttPred1))
confusionMatrix(table(DFTek$Attrition,kNN_Prediction))
# 249   2
# 40   9
#Sensitivity : 0.8616          
#Specificity : 0.8182          
#Pos Pred Value : 0.9920          
#Neg Pred Value : 0.1837

```


#Naive Bayes Prediction Model
```{r Naive Bayes prediction Model}
New_Naive_Bayes_Model4 <- naiveBayes(Attrition ~ OverTime + YearsAtCompany + YearsInCurrentRole + YearsWithCurrManager + JobLevel,data = DFTrain)
New_Naive_Bayes_Model4
New_NB_Prediction4 <- predict(New_Naive_Bayes_Model4, DFTrain[,c('OverTime','YearsAtCompany','YearsInCurrentRole','YearsWithCurrManager','JobLevel')])
confusionMatrix(table(New_NB_Prediction4,DFTrain$Attrition))
Test_Data_NB_Pred4 <- predict(New_Naive_Bayes_Model4, DFTest[,c('OverTime','YearsAtCompany','YearsInCurrentRole','YearsWithCurrManager','JobLevel')])
confusionMatrix(table(Test_Data_NB_Pred4,DFTest$Attrition))
#Test_Data_NB_Pred4  
#No Yes
# No  243  31
# Yes   8  18
# Accuracy : 0.87            
# 95% CI : (0.8266, 0.9059)
# No Information Rate : 0.8367          
# P-Value [Acc > NIR] : 0.065830        
# Kappa : 0.4136          
# Mcnemar's Test P-Value : 0.000427        
# Sensitivity : 0.9681          
# Specificity : 0.3673          
# Pos Pred Value : 0.8869          
# Neg Pred Value : 0.6923          
# Prevalence : 0.8367          
# Detection Rate : 0.8100          
# Detection Prevalence : 0.9133          
# Balanced Accuracy : 0.6677
```

###Recommendations Summary
####To recap the opportunities, 



#### Other References 
###Youtube PowerPoint Presentation
#### https://youtu.be/X7wWLbDmeDM
###Github Respository
####https://github.com/NicoleABartholow/MSDS6306CaseStudy2